---
title: "Coffee Shop Project"
author: "Nancy Wangare"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
Sys.which("make")
```

```{r}
options(repos=c(CRAN="https://cloud.r-project.org"))
```

## Packages

```{r}
pacman::p_load(
  tidyverse, tidylog, janitor, readxl, magrittr, skimr, ggplot2
)
```

## Load Data

```{r}
df <- read_excel("C:/Users/Nancy/Documents/Mysql project/coffee_shop_survey.xlsx")

```

```{r}

df_clean <- df %>%
  clean_names()

names(df_clean)
```

```{r}
skim(df_clean)
```

# Descriptive Analysis Questions

## 1.The average age of customers visiting the coffee shop

## 

```{r}
mean(df_clean$age)
```

## 2.Gender distribution among respondents

```{r}
gender_distribution <- df_clean  %>% 
  count(gender, name="count") 

  gender_distribution
```

```{r}
ggplot(gender_distribution, aes(x= gender, y=count,fill= gender))+
   geom_col()+
   labs(
     title= "Gender Distribution of Customers",
     y= "Total number of customers",
     x= "Gender of customers"
   )+
   theme_minimal()
```

## 3.The most and least popular products

```{r}
library(dplyr)
```

```{r}
product_popularity <- df_clean %>% 
  count(favorite_product, name =  "count") %>% 
  arrange(desc(count))
```

# Relationship & Comparison Questions

## 1. Does visit frequency differ by gender or age group

```{r}
aggregate(visit_frequency ~ gender, df_clean, FUN=mean)
```

```{r}
# Test for statistical significance:
kruskal.test(visit_frequency ~ gender, df_clean)
```

```{r}
# Creating Age Groups
df_clean$age_group <- cut(df_clean$age, breaks=c(0, 25, 40, 60, 100), labels=c("Young", "Middle-Aged", "Older", "Senior"))

```

```{r}
# Comparing visit frequency across age groups
aggregate(visit_frequency ~ age_group, df_clean, FUN=mean)
```

## 2.Is there a relationship between visit frequency and satisfaction score

##### Correlation

```{r}
cor(df_clean$visit_frequency, df_clean$satisfaction_score)
```

```{r}
ggplot(df_clean, aes(x = visit_frequency, y = satisfaction_score)) +
  geom_point() + 
  geom_smooth(method="lm", se=FALSE) + 
  theme_minimal() +
  labs(title="Visit Frequency vs Satisfaction Score",
       x="Visit Frequency", y="Satisfaction Score")
```

## 3.Do loyalty members spend more time at the coffee shop than non-members

```{r}
df_clean %>% 
  group_by(loyalty_member) %>% 
  summarise(avg_time_spent = mean(time_spent_min)) %>% 
  ggplot(aes(x=loyalty_member, y= avg_time_spent, fill = loyalty_member)) + 
  geom_col() +
  geom_text(aes(label = avg_time_spent, vjust = -0.5)) +
  theme_minimal()
```

##### Perform a t-test

```{r}
t.test(time_spent_min ~ loyalty_member, data = df_clean)
```

## 4.Are customers who would recommend the coffee shop more likely to be loyal members?

```{r}
chi_square_data <- table(df_clean$would_recommend, df_clean$loyalty_member)

print(chi_square_data)
```

```{r}
# Perform chi-square test
chisq.test(chi_square_data)
```

## Insightful/Advanced Questions

## 1.What factors (age, loyalty status, product preference) are associated with higher satisfaction scores

```{r}
df_model <- df_clean %>%
  select(age, loyalty_member, favorite_product, satisfaction_score) %>%
  mutate(
    loyalty_member = factor(loyalty_member),
    favorite_product = factor(favorite_product)
  )
```

```{r}
model <- lm(satisfaction_score ~ age + loyalty_member + favorite_product, data = df_model)
summary(model)
```

```{r}
library(broom)
```

```{r}
tidy(model) %>% 
  ggplot(aes(x = term, y = estimate, fill = p.value < 0.05)) +
  geom_col() +
  coord_flip() +
  labs(title = "Impact of Factors on Satisfaction Score",
       y = "Effect Size", x = "")
```

## 2.Do loyalty members spend more time at the coffee shop than non-members

```{r}
coffee_pastries <- df_clean %>% 
  filter(favorite_product %in% c('Coffee', 'Pastry')) %>% 
  group_by(favorite_product) %>% 
  summarise(average_satisfaction = mean(satisfaction_score)) %>% 
  ggplot(aes(x = favorite_product, y = average_satisfaction, fill = favorite_product)) +
  geom_col() +
  theme_minimal()
```

```{r}
print(coffee_pastries)
```

## 3.Are customers who would recommend the coffee shop more likely to be loyal members

```         
```

```{r}
df_model <- df_clean %>%   
  select(would_recommend, gender, satisfaction_score, age, loyalty_member,           visit_frequency) %>%   
  mutate(     recommend = ifelse(would_recommend == "Yes", 1, 0),         loyalty_member = as.factor(loyalty_member),gender = as.factor(gender)   ) %>%   drop_na()
```

##### Logistics Regression

```{r}
log_model <- glm(recommend ~ satisfaction_score + age + loyalty_member + visit_frequency + gender, data = df_model, family = "binomial")

summary(log_model)
```

```{r}

```
